---
title: "Keyness Analysis"
author: "Junhui Park"
date: "`r Sys.Date()`"
output: html_document

---

## Purpose
This file is to conduct a keyness analysis comparing a frequency between target corpus (post-period) and reference corpus (pre-period). Primarily, this is to identify terms that became relatively more prominent after the publication of the CCC policy recommendations.


## Workflow note
This document compares vocabulary in the pre- and post-CCC parliamentary climate-policy corpora using keyness statistics and complementary effect-size measures. It prepares the text, constructs document-feature matrices, filters and ranks candidate keywords, inspects contextual uses, and exports tables and figures. The transition year is excluded from the main pre/post comparison. Keyness describes differences in relative word use, not the causal effect of the CCC. Existing methodological notes and all code are preserved.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
# Load the packages and establish the options used by subsequent chunks. Reproducibility depends on the stated package versions, local input files and working environment.

library(xml2)
library(dplyr)
library(tidyverse)
library(stringr)
library(purrr)
library(tibble)
library(quanteda)
library(quanteda.textstats)
library(quanteda.textplots)
library(stopwords)
library(openxlsx)
library(ggrepel)

options(stringsAsFactors = FALSE)
set.seed(123)

knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE
)
```

---

# Data loading
```{r data loading, results='hide'}
# Read the parliamentary source files into R. The resulting objects provide the input for the extraction and filtering steps below.

all_files <- list.files(
  "xml_folder_path",
  pattern    = "\\.xml$",
  full.names = TRUE
)

length(all_files)
head(all_files, 5)
```

---

# XML parsing
```{r parsing}
# Extract speech text and relevant metadata from the XML structure, applying the cleaning operations specified in the code. Subsequent counts depend on the unit of text produced here.


parse_one_xml <- function(file_path) {
  doc <- read_xml(file_path)
  xml_ns_strip(doc)

  if (length(xml_find_all(doc, ".//paragraphe")) > 0) {
    nodes    <- xml_find_all(doc, ".//paragraphe")
    tag_used <- "paragraphe"
  } else if (length(xml_find_all(doc, ".//para")) > 0) {
    nodes    <- xml_find_all(doc, ".//para")
    tag_used <- "para"
  } else if (length(xml_find_all(doc, ".//texte")) > 0) {
    nodes    <- xml_find_all(doc, ".//texte")
    tag_used <- "texte"
  } else {
    return(tibble(
      file        = basename(file_path),
      tag_used    = NA_character_,
      speech_text = NA_character_
    ))
  }

  tibble(
    file        = basename(file_path),
    tag_used    = tag_used,
    speech_text = xml_text(nodes)
  ) %>%
    mutate(
      speech_text = str_remove_all(speech_text, "\\d+"),
      speech_text = str_remove_all(speech_text, "\\([^\\)]+\\)"),
      speech_text = str_remove_all(speech_text, "La parole est à.*?(\\.|$)"),
      speech_text = str_remove_all(speech_text, "La séance est.*?(\\.|$)"),
      speech_text = str_squish(speech_text)
    ) %>%
    filter(!is.na(speech_text), speech_text != "")
}
```

---

# Full corpus building
```{r full corpus building}
# Assemble the extracted speeches into the working corpus. Check the object structure and row counts before interpreting later subsets.


df_full <- map_dfr(all_files, parse_one_xml) %>%
  filter(!is.na(speech_text), speech_text != "")

dim(df_full)
table(df_full$tag_used, useNA = "ifany")
```

---

# Temporal variables
```{r year and period}
# Assign speeches to years and analytical periods. These labels define the denominators and comparison groups used later.

# NOTE: Pre period = 3 years (2017-2019); Post period = 2 years (2021-2022).
# This corpus-size imbalance is handled by G² (relative frequency based), but should be noted explicitly in the methods section.

df_full <- df_full %>%
 mutate(
  year = stringr::str_extract(file, "20\\d{2}") %>%
   as.numeric(),
  period = case_when(
   year %in% 2017:2019 ~ "pre",
   year == 2020 ~ "transition",
   year %in% 2021:2022 ~ "post",
   TRUE ~ NA_character_
 )
) %>%
 filter(year %in% 2017:2022)
```

---

# Dictionaries

## Climate-policy dictionary
```{r climate-policy dictionary}
# Define the general climate-policy terms used to identify the relevant parliamentary speech subset. This dictionary defines the analytical corpus, not explicit CCC references.

climate_keywords <- c(
  "climat", "climatique",
  "environnement", "écologie",
  "transition écologique",
  "transition énergétique",
  "émissions",
  "carbone",
  "pollution",
  "fiscalité carbone",
  "pesticides",
  "déchets",
  "recyclage",
  "passoires énergétiques",
  "neutralité carbone",
  "empreinte carbone",
  "énergies renouvelables",
  "sobriété énergétique",
  "rénovation énergétique",
  "efficacité énergétique",
  "économie circulaire",
  "zéro artificialisation nette",
  "transition juste",
  "report modal",
  "covoiturage",
  "alimentation durable",
  "engrais minéraux azotés",
  "sortie du pétrole"
)
```

## CCC explicit dictionary
```{r CCC explicit dictionary}
# Define, combine or inspect the keyword set used in the subsequent matching step. Keyword selection determines the breadth of the resulting indicator.

ccc_explicit <- c(
  "convention citoyenne",
  "convention citoyenne pour le climat",
  "convention citoyenne sur le climat",
  "conventions citoyennes",
  "citoyens de la convention",
  "citoyens tirés au sort",
  "150 citoyens"
)
```

## CCC policy-theme dictionary
```{r CCC policy-theme dictionary}
# Define, combine or inspect the keyword set used in the subsequent matching step. Keyword selection determines the breadth of the resulting indicator.

consumption <- c(
  "surconsommation", "emballage", "plastique à usage unique",
  "publicité", "suremballage", "consommation responsable",
  "affichage environnemental", "affichage carbone",
  "écoresponsables", "écoresponsabilité", "produits polluants",
  "produits vertueux", "recyclage", "économie circulaire",
  "déchets d'emballage", "utilisation du plastique", "vrac",
  "réutilisable", "contenants réutilisables", "bouteilles plastiques",
  "réduction des déchets", "matériaux biodégradables", "réemploi"
)

production_work <- c(
  "produits durables", "écoconception", "longévité des produits",
  "investissements verts", "clauses environnementales",
  "droit pollueur-payeur", "filières émettrices", "écocide",
  "produits micropolluants", "énergie verte", "énergies renouvelables",
  "méthanisation", "autoconsommation", "écolabels", "énergie décarbonée"
)

transport_mobility <- c(
  "mobilité durable", "véhicules faibles émissions",
  "mobilités douces", "véhicules partagés", "mobilités alternatives",
  "zones de faible émission", "pistes cyclables",
  "véhicules électriques", "véhicules hybrides",
  "émissions du véhicule", "écoconduite", "multimodalité",
  "moteur éthanol", "moteur électrique", "moteur thermique",
  "covoiturage", "biocarburants"
)

housing <- c(
  "construction bas carbone", "performance énergétique",
  "rénovation globale", "chaudières au fioul",
  "passoires énergétiques", "bâtiments énergivores",
  "audit énergétique", "certificats d'économies d'énergie",
  "déperditions de chaleur", "climatisation", "chauffage au fioul"
)

food <- c(
  "alimentaire durable", "EGalim", "repas végétariens",
  "restauration collective", "faible coût environnemental",
  "transition vers le bio", "gaspillage alimentaire",
  "pratiques agroécologiques", "agroécologie",
  "réduction des pesticides", "agriculture biologique",
  "plan protéagineux", "prairies permanentes",
  "reconversion de l'élevage", "développement des légumineuses",
  "rediversifier les cultures", "engrais chimiques", "couvert végétal",
  "élevages de ruminants", "ecoscheme", "pêche durable", "surpêche",
  "propulsion verte", "clauses environnementales",
  "consommation de viande", "régime moins carné"
)

constitution_gov <- c(
  "crime d'écocide", "érosion de la biodiversité",
  "utilisation des sols", "pollution chimique",
  "dommage écologique", "préjudice écologique"
)

ccc_policy <- unique(c(
  consumption, production_work, transport_mobility,
  housing, food, constitution_gov
))
```

---

# Corpus filtering

## Climate-policy corpus
```{r climate-corpus}
# Construct or inspect the working speech corpus and its analytical subsets.


climate_pattern <- regex(
 paste0("\\b(", paste(climate_keywords, collapse = "|"), ")\\b"), ignore_case = TRUE
)

df_climate <- df_full %>%
 filter(str_detect(speech_text, climate_pattern))

dim(df_climate)
nrow(df_climate) / nrow(df_full)

df_climate %>%
 count(period) %>%
 mutate(ratio = n / sum(n))
```


```{r climate-sample, results='hide'}
# Construct or inspect the working speech corpus and its analytical subsets.

set.seed(123)
df_climate %>%
  slice_sample(n = 10) %>%
  select(year, period, speech_text)
```

## CCC explicit corpus
```{r ccc-explicit-corpus}
# Construct or inspect the working speech corpus and its analytical subsets.


explicit_ccc_pattern <- regex(
 paste0(
  "\\b(",
  paste(ccc_explicit, collapse = "|"),
  ")\\b"
 ),
 ignore_case = TRUE
)

df_ccc_explicit <- df_climate %>%
 filter(
  str_detect(
   speech_text,
   explicit_ccc_pattern
  )
 )

dim(df_ccc_explicit)
nrow(df_ccc_explicit) / nrow(df_climate)
table(df_ccc_explicit$period, useNA = "ifany")

df_ccc_explicit %>%
 count(period) %>%
 mutate(ratio = n / sum(n))
```

## CCC policy-theme corpus
```{r ccc-policy-corpus}
# Construct or inspect the working speech corpus and its analytical subsets.


ccc_policy_pattern <- regex(
 paste0(
  "\\b(",
  paste(ccc_policy, collapse = "|"),
  ")\\b"),
ignore_case = TRUE)

df_ccc_policy <- df_climate %>%
 filter(
  str_detect(
   speech_text,
   ccc_policy_pattern
  )
 )

ccc_policy_period_ratio <- df_ccc_policy %>%
  count(period, name = "ccc_policy_count") %>%
  left_join(
    df_climate %>% count(period, name = "climate_total"),
    by = "period"
  ) %>%
  mutate(
    proportion = ccc_policy_count / climate_total,
    percentage = round(proportion * 100, 2),
    proportion = round(proportion, 3),
    period     = factor(period, levels = c("pre", "transition", "post"))
  ) %>%
  arrange(period)

ccc_policy_period_ratio
```

---

# Keyness analysis
The study employs the R quanteda package (Benoit et al. 2018). The analysis is conducted with three steps: 1) corpus preparation, 2) keyness statistical analysis, and 3) peak keywords selection based on KWIC and iterative manual inspection of sample texts.

## Prepare corpus (pre vs post only)
```{r keyness-prep}
# Prepare the pre/post comparison corpus, excluding the transition year as specified. The two periods contain different numbers of calendar years.

# Transition year (2020) excluded from main comparison.
# Apostrophes are removed entirely ("") rather than replaced with a space (" ") to prevent words like "c'est" from being split into two tokens ("c", "est"), which would leave residual single-character tokens that are difficult to remove via stopwords.

df_keyness <- df_climate %>%
  filter(period %in% c("pre", "post")) %>%
  mutate(
    comparison_period = period,
    speech_text_clean = speech_text %>%
     str_to_lower() %>%
     str_replace_all("\\b(l|d|j|n|m|s|c|qu)'", " ") %>%
     str_replace_all("\\b(l|d|j|n|m|s|c|qu)\u2019", " ") %>%
     str_replace_all("[''\u2019]", "") %>%
     str_squish()
  )

table(df_keyness$comparison_period)
```

## Tokenisation
```{r tokenisation}
# Convert texts into tokens and apply the stated token-processing choices before constructing the document-feature matrix.

corp_keyness <- corpus(df_keyness, text_field = "speech_text_clean")
docvars(corp_keyness, "comparison_period") <- df_keyness$comparison_period

toks_keyness <- tokens(
  corp_keyness,
  remove_punct   = TRUE,
  remove_numbers = TRUE,
  remove_symbols = TRUE
) %>%
  tokens_tolower()

# Compound multi-word expressions
# French parliamentary discourse contains many meaningful compound terms.
# Without compounding, "transition écologique" is split into two separate unigrams, diluting its analytical signal. tokens_compound() preserves these as single features (e.g., transition_écologique).

mwe_list <- phrase(c(
  # Climate policy general
  "transition écologique", "transition énergétique",
  "fiscalité carbone", "neutralité carbone", "empreinte carbone",
  "énergies renouvelables", "sobriété énergétique",
  "rénovation énergétique", "efficacité énergétique",
  "économie circulaire", "zéro artificialisation nette",
  "transition juste", "report modal",
  "alimentation durable", "engrais minéraux azotés", "sortie du pétrole",
  # CCC explicit
  "convention citoyenne", "convention citoyenne pour le climat",
  "conventions citoyennes", 
  # Consumption
  "plastique à usage unique", "consommation responsable",
  "affichage environnemental", "affichage carbone",
  "déchets demballage", "réduction des déchets",
  "matériaux biodégradables", "contenants réutilisables",
  "bouteilles plastiques",
  # Production
  "produits durables", "longévité des produits",
  "investissements verts", "clauses environnementales",
  "droit pollueur-payeur", "filières émettrices",
  "énergie décarbonée", "énergie verte",
  # Transport
  "mobilité durable", "véhicules faibles émissions",
  "mobilités douces", "véhicules partagés", "mobilités alternatives",
  "zones de faible émission", "pistes cyclables",
  "véhicules électriques", "véhicules hybrides",
  "émissions du véhicule",
  # Housing
  "construction bas carbone", "performance énergétique",
  "rénovation globale", "chaudières au fioul",
  "passoires énergétiques", "bâtiments énergivores",
  "audit énergétique", "déperditions de chaleur", "chauffage au fioul",
  # Food
  "repas végétariens", "restauration collective",
  "gaspillage alimentaire", "pratiques agroécologiques",
  "réduction des pesticides", "agriculture biologique",
  "plan protéagineux", "prairies permanentes",
  "reconversion de lélevage", "développement des légumineuses",
  "engrais chimiques", "couvert végétal", "élevages de ruminants",
  "pêche durable", "consommation de viande", "régime moins carné",
  # Constitution
  "crime décocide", "érosion de la biodiversité",
  "utilisation des sols", "pollution chimique",
  "dommage écologique", "préjudice écologique"
))

toks_keyness <- tokens_compound(toks_keyness, pattern = mwe_list)

# Stopwords removal
# Note: apostrophes have been removed entirely in the pre-processing step,
# so contractions like "c'est" become "cest", "d'un" becomes "dun", etc.
# These collapsed forms are included explicitly in the stopword list below.
toks_keyness <- tokens_remove(
  toks_keyness,
  pattern = c(
    stopwords("fr"),
    "mme", "monsieur", "madame", "ministre",
    "président", "présidente", "collègue", "collègues",
    "rapporteur", "rapporteure", "député", "députés",
    "séance", "commission", "article", "articles", "assemblée",
    "l", "d", "j", "n", "m", "s", "c",
    "être", "donc", "très", "comme", "tout", "aussi",
    "bien", "si", "tous", "plus", "non", "a",
    # noise-words removed via KWIC and sample text inspection
    "délégué", "déléguée", "délégués", "déléguées",
    "soutien", "aides", "démocrates", "harcèlement",
    "barbara", "julien", "chantal", "cendra",
    "cnr", "traversons"
  )
)

# NOTE on lemmatisation:
# French morphology means inflected forms (e.g., "rénovation"/"rénovations", "climatique"/"climatiques") are counted as distinct features.
# Lemmatisation via the {udpipe} package with a French UD model would reduce this noise but was not applied here due to computational cost.

```

## Build DFM
```{r build-dfm, results='hide'}
# Construct the document-feature matrix and apply the stated frequency filtering prior to the keyness comparison.

dfm_keyness <- dfm(toks_keyness)

# Global minimum frequency filter (applied before grouping): removes features appearing fewer than 5 times across the full corpus.
# This is a conservative pre-filter; a stricter threshold (total >= 10) is applied after grouping in the low-frequency filter step below.

dfm_keyness <- dfm_trim(dfm_keyness, min_termfreq = 5)

topfeatures(dfm_keyness, 30)
dfm_keyness

# Group by period for keyness comparison
dfm_grouped <- dfm_group(dfm_keyness, groups = comparison_period)
```

## Run keyness (G²)
```{r run-keyness}
# Calculate the G² keyness statistic to identify vocabulary whose relative frequency differs between the comparison periods.

# G² (log-likelihood ratio) — primary statistic.
# Preferred over chi² for corpus linguistics: more robust with low-frequency cells.
# keyness_lr_raw preserves the original "keyness" class required by textplot_keyness(). Renaming or relocating columns strips this class, so a separate working object (keyness_lr) is created for downstream analysis.

keyness_lr_raw <- textstat_keyness(
  dfm_grouped,
  target  = "post",
  measure = "lr"
)

keyness_lr <- keyness_lr_raw %>%
  rename(pre = "n_reference", post = "n_target") %>%
  relocate(pre:post, .after = "feature")
```

## Effect size: Log Ratio + 95% CI (Hardie 2014)
```{r effect-size}
# Log Ratio measures the relative frequency difference between post and pre corpora on a log2 scale (LR = 1 means twice as frequent in post; LR = -1 means half).
# 0.5 smoothing is applied to handle zero counts (standard practice).
# 95% CI is computed following Hardie (2014): if the CI crosses zero, the direction of the effect is uncertain and the keyword is flagged unreliable.


total_pre  <- sum(dfm_grouped["pre",  ])
total_post <- sum(dfm_grouped["post", ])

keyness_ef <- keyness_lr %>%
  mutate(
    post_r    = (post + 0.5) / total_post,
    pre_r     = (pre  + 0.5) / total_pre,
    log_ratio = log2(post_r / pre_r),
    margin    = 1.96 * sqrt(1 / post + 1 / pre),
    ci_lower  = log_ratio - margin,
    ci_upper  = log_ratio + margin,
    reliable  = (ci_lower > 0 | ci_upper < 0)
  ) %>%
  select(-post_r, -pre_r, -margin)
```

## Multiple testing correction (Benjamini-Hochberg)
```{r multiple-testing}
# BH controls the False Discovery Rate rather than the Family-Wise Error Rate (Bonferroni). This is the appropriate choice for exploratory corpus analysis where a large number of features are tested simultaneously and some false positives are acceptable in exchange for higher statistical power.

keyness_ef <- keyness_ef %>%
  mutate(p_G2_adj = p.adjust(p, method = "BH")) %>%
  relocate(p_G2_adj, .after = "p")
```

## Low-frequency inspection and filter
```{r low-freq-inspect, results='hide'}
# Step 1: Inspect low-frequency items before filtering.
# These are items where effect size estimates are unreliable regardless of G² significance (small absolute counts amplify noise).

low_freq_items <- keyness_ef %>%
  mutate(total_count = pre + post) %>%
  filter(total_count < 10) %>%
  arrange(desc(log_ratio))

cat("Number of features with total count < 10:", nrow(low_freq_items), "\n")
print(low_freq_items %>% select(feature, pre, post, G2, log_ratio))
```

```{r low-freq-filter, results='hide'}
# Step 2: Apply the dual filter for reliable final keyword list.
# All four criteria must be met simultaneously:
#   (a) total_count >= 10  — removes unstable low-count estimates
#   (b) p_G2_adj < 0.05   — statistically significant after BH correction
#   (c) |log_ratio| >= 1  — at least 2x relative frequency difference
#   (d) reliable == TRUE  — 95% CI does not cross zero (Hardie 2014)

keyness_reliable <- keyness_ef %>%
  mutate(total_count = pre + post) %>%
  filter(
    total_count    >= 10,
    p_G2_adj       <  0.05,
    abs(log_ratio) >= 1,
    reliable       == TRUE
  ) %>%
  arrange(desc(G2))

cat("Number of reliable keywords after dual filter:", nrow(keyness_reliable), "\n")

```

---

# Final keyword tables

## Top 50 by G² (statistical salience)
```{r top50-g2, results='hide'}
# ci_lower and ci_upper are omitted: all features in keyness_reliable have already passed the reliable == TRUE filter (CI does not cross zero).

top50_G2 <- keyness_reliable %>%
  arrange(desc(G2)) %>%
  select(feature, pre, post, G2, p_G2_adj, log_ratio) %>%
  head(50)

top50_G2
```

## Top 50 by Log Ratio (effect size)
```{r top50-lr, results='hide'}
# Calculate or display the relevant period-level proportions. The interpretation depends on the numerator, denominator and dictionary specification used in this chunk.


top50_lr <- keyness_reliable %>%
  arrange(desc(log_ratio)) %>%
  select(feature, pre, post, log_ratio, G2, p_G2_adj) %>%
  head(50)

top50_lr
```

## Top 50 negative keyness (pre > post)
```{r top50-negative, results='hide'}
# Carry out the 'Top 50 negative keyness (pre > post)' step and inspect its output before proceeding. This chunk forms part of the documented workflow; its results have not been independently re-run here.

# Words significantly more frequent in the pre-CCC period.

top50_negative <- keyness_ef %>%
  mutate(total_count = pre + post) %>%
  filter(
    total_count    >= 10,
    p_G2_adj       <  0.05,
    log_ratio      <= -1,
    reliable       == TRUE
  ) %>%
  arrange(log_ratio) %>%
  select(feature, pre, post, log_ratio, G2, p_G2_adj) %>%
  head(50)

top50_negative
```

---

# Visualisation
```{r plot-keyness, fig.width=10, fig.height=8}
# Produce the figure for the preceding analysis. Check the axis labels, denominator and time grouping against the data object used in this chunk.

# Plot showing top 15 reliable keywords by G2

top15_G2 <- top50_G2 %>%
 slice_max(G2, n = 15) %>%
 arrange(G2)

ggplot(
 top15_G2,
 aes(
  x = reorder(feature, G2),
  y = G2
 )
) +
 geom_col() +
 coord_flip() +
 labs(
  title = expression("Top 15 reliable keywords by " * G^2),
  x = NULL,
  y = expression(G^2)
 ) +
 theme_minimal(base_size = 12)


ggsave(
 "Top15_G2_pre_post.png",
 width = 8, height = 6, dpi = 300
)

```


```{r plot-logratio, fig.width=10, fig.height=7}
# Produce the figure for the preceding analysis. Check the axis labels, denominator and time grouping against the data object used in this chunk.
# Log Ratio plot with 95% CI for top 20 reliable keywords.
# Point position = Log Ratio; error bars = 95% CI (Hardie 2014).
# Dotted vertical lines mark the |LR| = 1 threshold.

top20_plot <- keyness_reliable %>%
  arrange(desc(abs(log_ratio))) %>%
  head(20) %>%
  mutate(
    direction = if_else(log_ratio > 0,
                        "post (more frequent)",
                        "pre (more frequent)"),
    feature   = reorder(feature, log_ratio)
  )

ggplot(top20_plot, aes(x = log_ratio, y = feature, colour = direction)) +
  geom_point(size = 2.5) +
  geom_errorbarh(
    aes(xmin = ci_lower, xmax = ci_upper),
    height = 0.3, linewidth = 0.6
  ) +
  geom_vline(xintercept = 0,
             linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = c(-1, 1),
             linetype = "dotted", colour = "grey70") +
  scale_colour_manual(
    values = c("post (more frequent)" = "#2166ac",
               "pre (more frequent)"  = "#d6604d")
  ) +
  labs(
    title    = "Top 20 distinctive keywords by Log Ratio (pre vs post-CCC)",
    x        = "Log Ratio (log2 scale)",
    y        = NULL,
    colour   = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

ggsave(
 "Top20_LR_pre_post.png",
 width = 8, height = 6, dpi = 300
)

```

---

# Export results
```{r export, results='hide'}
# Write the analysis outputs to the file paths specified in the code. Confirm those paths exist in the execution environment before re-running.

summary_stats <- tibble(
  statistic = c("Min", "Q1", "Median", "Mean", "Q3", "Max"),
  G2        = as.numeric(summary(keyness_reliable$G2)),
  log_ratio = as.numeric(summary(keyness_reliable$log_ratio))
) %>%
  mutate(across(where(is.numeric), ~ round(., 3)))

write.xlsx(
  list(
    summary_stats  = as.data.frame(summary_stats),
    top50_G2       = as.data.frame(top50_G2),
    top50_lr       = as.data.frame(top50_lr),
    top50_negative = as.data.frame(top50_negative),
    all_reliable   = as.data.frame(keyness_reliable),
    all_combined   = as.data.frame(keyness_ef)
  ),
  file = "keyness_results_final.xlsx"
)

cat("Export complete.\n")
```

---

# KWIC inspection

## Keyword selection
```{r kwic-selection-export}
# Export keyness_reliable sorted by G² and Log Ratio separately for manual inspection in Excel. Keywords appearing in the upper range of both distributions are prioritised for KWIC inspection.

write.xlsx(
  list(
    by_G2        = keyness_reliable %>%
                     arrange(desc(G2)) %>%
                     select(feature, pre, post, G2, p_G2_adj, log_ratio) %>%
     mutate(
      G2 = round(G2, 2),
      log_ratio = round(log_ratio, 2),
      p_G2_adj = case_when(
       p_G2_adj < 0.001 ~ "< .001",
       TRUE ~ as.character(round(p_G2_adj, 3))
      )
     ),
    by_log_ratio = keyness_reliable %>%
                     arrange(desc(log_ratio)) %>%
                     select(feature, pre, post, G2, p_G2_adj, log_ratio) %>%
     mutate(
      G2 = round(G2, 2),
      log_ratio = round(log_ratio, 2),
      p_G2_adj = case_when(
       p_G2_adj < 0.001 ~ "< .001",
       TRUE ~ as.character(round(p_G2_adj, 3))
      )
     )
  ),
  file = "keyness_inspection_final.xlsx"
)
```

## KWIC concordance lines
```{r kwic-concordance}
# Concordance lines (window = 10 tokens) for keywords selected from the inspection spreadsheet. Update the pattern list based on the Excel review.
# Priority: keywords with ambiguous cluster assignment, potential noise, and those appearing in the upper range of both G² and Log Ratio.

kwic_keywords <- kwic(
  toks_keyness,
  pattern = c(
    "relance", "crise", "plan", "présidence",
    "aéronautique", "publicité", "textile", "animaux",
    "chargé"
  ),
  window    = 10
) %>%
  as.data.frame() %>%
 mutate(docname_num = str_extract(docname, "\\d+")) %>%
  left_join(
    df_keyness %>%
      mutate(docname_num = as.character(row_number())) %>%
      select(docname_num, year, period),
    by = "docname_num"
  ) %>%
 select(-docname_num)

write.xlsx(kwic_keywords, "kwic_keywords_final.xlsx")

# window enlargement
kwic_chargé_w20 <- kwic(
  toks_keyness,
  pattern = 
    "chargé",
  window    = 20
) %>%
  as.data.frame() %>%
 mutate(docname_num = str_extract(docname, "\\d+")) %>%
  left_join(
    df_keyness %>%
      mutate(docname_num = as.character(row_number())) %>%
      select(docname_num, year, period),
    by = "docname_num"
  ) %>%
 select(-docname_num)

write.xlsx(kwic_chargé_w20, "kwic_chargé_w20.xlsx")

```


# sample text inspection
```{r text inspection}
# Inspect intermediate results for coverage, plausible matches and potential noise. This is a diagnostic step rather than a separate inferential test.

df_keyness %>%
  filter(str_detect(speech_text,
                    regex(
"\\b(présidence|traversons|chargé|cnr|textile|animaux)\\b", 
ignore_case = TRUE))) %>%
 mutate(keyword = str_extract(speech_text,
"\\b(présidence|traversons|chargé|CNR|textile|animaux)\\b")) %>%
 group_by(keyword, year, period) %>%
 slice_sample(n = 5) %>%
 select(keyword, year, period, speech_text) %>%
 arrange(keyword, year) %>%
 write.csv("sample_inspection_newly_added.csv", row.names = FALSE, fileEncoding = "UTF-8")

# inspection per word
plan_sample <- df_keyness %>%
  filter(str_detect(speech_text,
                    regex("\\bplan\\b", ignore_case = TRUE))) %>%
 mutate(keyword = str_extract(speech_text,
                              regex("\\bplan\\b", ignore_case = TRUE))
) %>%
 group_by(year, period) %>%
 slice_sample(prop = 1) %>%
 slice_head(n = 5) %>%
 ungroup() %>%
 select(keyword, year, period, speech_text) %>%
 arrange(year, period)

 write.xlsx(plan_sample, "sample_inspection_plan.xlsx")
```

---

# corpus size
```{r}
# Construct or inspect the working speech corpus and its analytical subsets.

corpus_size_yearly <- df_keyness %>%
 mutate(
  n_words = str_count(speech_text, "\\S+")
 ) %>%
 group_by(year) %>%
 summarise(
  speeches = n(),
  total_words = sum(n_words),
  .groups = "drop"
 )

corpus_size_yearly

```

---

# Peaks in keywords
```{r peak in keywords}
# Count frequency of each peak keyword per year.
# Keywords are lowercased to avoid duplicate counting (e.g. "Présidence" vs "présidence").

peak_keywords <- c(
 "relance", "crise", "résilience",
 "convention citoyenne pour le climat",
 "dérèglement", "aéronautique",
 "néonicotinoïdes","écocide"
)

# yearly peak keywords
peak_yearly <- df_keyness %>%
 select(year, period, speech_text) %>%
 mutate(
  speech_text_lower = str_to_lower(speech_text)
 ) %>%
 crossing(keyword = peak_keywords) %>%
  filter(
   str_detect(speech_text, regex(paste0("\\b", keyword, "\\b"))
  )
   ) %>%
  count(keyword, year, name = "n")

# Export wide-format table for reporting
peak_yearly_wide <- peak_yearly %>%
  pivot_wider(
    names_from  = year,
    values_from = n,
    values_fill = 0
  ) %>%
  arrange(keyword)

write.xlsx(peak_yearly_wide, "peak_keywords_yearly.xlsx")

# normalised yearly peak keywords
peak_yearly_norm <- peak_yearly %>%
 left_join(corpus_size_yearly, by = "year") %>%
 mutate(
  freq_100k = (n / total_words) * 100000
 )

# Export wide-format table for normalised version
peak_year_norm_wide <- peak_yearly_norm %>%
 select(keyword, year, freq_100k) %>%
 pivot_wider(
  names_from = year,
  values_from = freq_100k,
  values_fill = 0
 ) %>%
 arrange(keyword)

peak_year_norm_wide %>%
 mutate(across(where(is.numeric), ~ round(., 2))) %>%
write.xlsx("peak_keywords_yearly_norm.xlsx")


```

---

# Visualisation
```{r plot-main, fig.width=10, fig.height=6}
# Produce the figure for the preceding analysis. Check the axis labels, denominator and time grouping against the data object used in this chunk.

# Main text figure: top 8 keywords by G² (directly specified from keyness_reliable)

keyword_order <- peak_yearly %>%
 group_by(keyword) %>%
 summarise(total_n = sum(n, na.rm = TRUE),
           .groups = "drop") %>%
 arrange(desc(total_n)) %>%
 pull(keyword)

peak_yearly <- peak_yearly %>%
 mutate(
  keyword = factor(keyword, levels = keyword_order)
 )

yearly_trends_raw <- peak_yearly %>%
 ggplot(aes(
  x = year,
  y = n,
  colour = keyword, group = keyword
 )) +
 geom_line(linewidth = 0.8) +
 geom_point(size = 2) +
 geom_vline(
  xintercept = 2020,
  linetype = "dashed",
  colour = "grey50"
 ) +
 annotate(
  "text",
  x = 2020.1,
  y = Inf,
  label = "CCC publication",
  hjust = 0,
  vjust = 1.5,
  size = 3.5,
  colour = "grey50"
 ) +
 scale_x_continuous(
  breaks = 2017:2022
 ) +
 labs(
  title = "Yearly frequency of selected peak keywords",
  subtitle = "Raw counts within the climate-policy corpus",
  x = NULL,
  y = "Frequency",
  colour = "Keyword"
 ) +
 theme_minimal(base_size = 12) +
 theme(
  legend.position = "right",
  legend.title = element_text(face = "bold"),
  plot.title = element_text(face = "bold"),
  plot.subtitle = element_text(size = 10)
 )

yearly_trends_raw

ggsave("yearly_trends_raw.png", yearly_trends_raw, width = 10, height = 6, dpi = 300)

# normalised frequency (robustness)
keyword_order <- peak_yearly_norm %>%
 group_by(keyword) %>%
 summarise(total_freq = sum(freq_100k, na.rm = TRUE),
.groups = "drop") %>%
 arrange(desc(total_freq)) %>%
 pull(keyword)

yearly_trends_norm <- peak_yearly_norm %>%
 ggplot(aes(
  x = year,
  y = freq_100k,
  colour = keyword,
  group = keyword
 )
 ) +
 geom_line(linewidth = 0.8) +
 geom_point(size = 2) +
 
 geom_vline(
  xintercept = 2020,
  linetype = "dashed",
  colour = "grey50"
 ) +
 annotate(
  "text",
  x = 2020.1,
  y = Inf,
  label = "CCC publication",
  hjust = 0,
  vjust = 1.5,
  size = 3.5,
  colour = "grey50"
 ) +
 scale_x_continuous(
  breaks = 2017:2022
 ) +
 labs(
  title = "Normalised yearly frequency of selected peak keywords",
  subtitle = "Frequency per 100,000 words within the climate-policy corpus",
  x = NULL,
  y = "Frequency per 100,000 words",
  colour = "Keyword"
 ) +
 theme_minimal(base_size = 12) +
 theme(
  legend.position = "right",
  legend.title = element_text(face = "bold"),
  plot.title = element_text(face = "bold"),
  plot.subtitle = element_text(size = 10)
 )

yearly_trends_norm

ggsave("yearly_trends_norm.png", yearly_trends_norm, width = 10, height = 6, dpi = 300)

```


---

# Methodological notes (for methods section)

1. **Corpus size imbalance**: The pre-CCC period spans three years (2017-2019) while the post-CCC period covers two years (2021-2022). G² operates on normalised relative frequencies and therefore accounts for this difference in corpus size. Nevertheless, the 2017 sub-corpus is substantially smaller than subsequent years, which may introduce instability in pre-period estimates.

2. **G² over chi²**: G² (log-likelihood ratio) is preferred over chi² for corpus keyness analysis as it is more robust when observed or expected cell frequencies are low, a common occurrence in large-scale lexical comparisons (Gabrielatos 2018).

3. **Multiple testing correction**: Given that keyness analysis involves simultaneous testing across a large number of features, p-values are adjusted using the Benjamini-Hochberg (BH) procedure, which controls the False Discovery Rate. Bonferroni correction was considered but rejected as overly conservative for exploratory corpus analysis (Benjamini and Hochberg 1995).

4. **Effect size threshold**: A Log Ratio threshold of |LR| >= 1 (equivalent to a minimum twofold difference in relative frequency) is applied in combination with statistical significance to distinguish substantively meaningful keywords from statistically significant but practically negligible differences (Hardie 2014).

5. **Confidence intervals for Log Ratio**: Following Hardie (2014), 95% confidence intervals are computed for all Log Ratio estimates. Features whose CI crosses zero are excluded from the final keyword list regardless of G² significance, as the direction of the effect cannot be considered reliable.

6. **Lemmatisation**: French morphological inflection means that related surface forms are counted as distinct features. Lemmatisation was not applied in this analysis, which constitutes a methodological limitation to be acknowledged.

7. **Multi-word expressions**: Key compound terms (e.g., transition ecologique, convention citoyenne) are compounded prior to tokenisation using tokens_compound() to preserve their lexical integrity as single analytical units.
