---
title: "cluster-level robustness check"
author: "Junhui Park"
date: "`r Sys.Date()`"
output: html_document
---

## Input and corpus version
The main analyses use the climate-policy corpus constructed in "1_full corpus building_final.Rmd".
In the analysis files, 'df_climate' refers to the main climate-policy corpus unless otherwise stated.
A stricter matching strategy using word boundaries is used only for validation or robustness checks.

## Robustness specification
This file uses a stricter matching strategy and/or stricter dictionary specification.
These results are used as robustness checks and do not replace the main analysis.

## Workflow note
This document examines the robustness of the policy-theme salience results at the level of the six CCC policy clusters. It reconstructs the speech corpus, applies cluster-specific keyword sets, checks selected matches, computes cluster-level counts and period proportions, and visualises the patterns. Cluster matches indicate thematic references, which need not be explicit references to the CCC. All original code is preserved.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)

# Load the packages and establish the options used by subsequent chunks. Reproducibility depends on the stated package versions, local input files and working environment.

# libraries
library(xml2)
library(dplyr)
library(tidyverse)
library(stringr)
library(purrr)
library(tibble)
library(ggplot2)

# global options
options(stringsAsFactors = FALSE)

# reproductibility
set.seed(123)

# knitr options
knitr::opts_chunk$set(
 echo = TRUE,
 warning = FALSE,
 message = FALSE)
```

## loading data
```{r data}
# Load the input data and construct the objects required by subsequent processing steps.

all_files <- list.files("xml_folder_path",
                    pattern = "\\.xml$",
                    full.names = TRUE)

length(all_files)
head(all_files, 5)

test_file <- all_files[1]
test_file

```

## parsing
```{r parsing, results='hide'}
# Extract and standardise text and metadata from the source records for the subsequent analysis.

parse_one_xml <- function(file_path) {
 doc <- read_xml(file_path)
 xml_ns_strip(doc)
 
 if (length(xml_find_all(doc, ".//paragraphe")) > 0) {
  nodes <- xml_find_all(doc, ".//paragraphe")
  tag_used <- "paragraphe"
 } else if (length(xml_find_all(doc, ".//para")) > 0) {
  nodes <- xml_find_all(doc, ".//para")
  tag_used <- "para"
 } else if (length(xml_find_all(doc, ".//texte")) > 0) {
  nodes <- xml_find_all(doc, ".//texte")
  tag_used <- "texte"
 } else {
  return(tibble(
         file = basename(file_path),
         tag_used = NA_character_,
         speech_text = NA_character_
         ))
 }
 
tibble(
 file = basename(file_path),
 tag_used = tag_used,
 speech_text = xml_text(nodes)
) %>%
 mutate(
  speech_text = str_remove_all(speech_text, "\\d+"),
  speech_text = str_remove_all(speech_text, "\\([^\\)]+\\)"),
  speech_text = str_remove_all(speech_text, "La parole est à.*?(\\.|$)"),
  speech_text = str_remove_all(speech_text, "La séance est.*?(\\.|$)"),
  speech_text = str_squish(speech_text)
 ) %>%
 filter(!is.na(speech_text), speech_text != "")
}

```

## full corpus building
```{r corpus building, results='hide'}
# Construct or inspect the working speech corpus and its analytical subsets.

df_full <- map_dfr(all_files, parse_one_xml)

dim(df_full)
head(df_full, 10)
table(df_full$tag_used, useNA = "ifany")

df_full <- df_full %>%
 filter(!is.na(speech_text), speech_text != "")
```

## add year and period
```{r periodisation}
# Calculate or display the year-specific results to assess temporal variation beyond the aggregated period comparison.

df_full <- df_full %>%
 mutate(
  year = stringr::str_extract(file, "20\\d{2}") %>%
   as.numeric(),
  period = case_when(
   year %in% 2017:2019 ~ "pre",
   year == 2020 ~ "transition",
   year %in% 2021:2022 ~ "post",
   TRUE ~ NA_character_
 )
) %>%
 filter(year %in% 2017:2022)

table(df_full$period, useNA = "ifany")

dim(df_full)
table(df_full$year)
table(df_full$period, useNA = "ifany")
```

## climate keywords
```{r climate keywords}
# Define, combine or inspect the keyword set used in the subsequent matching step. Keyword selection determines the breadth of the resulting indicator.

climate_keywords <- c(
 "climat", "climatique",
 "environnement", "écologie",
 "transition écologique",
 "transition énergétique",
 "émissions",
 "carbone",
 "pollution",
 "fiscalité carbone",
 "pesticides",
 "déchets",
 "recyclage",
 "passoires énergétiques",
 "neutralité carbone",
 "empreinte carbone",
 "énergies renouvelables",
 "sobriété énergétique",
 "rénovation énergétique",
 "efficacité énergétique",
 "économie circulaire",
 "zéro artificialisation nette",
 "transition juste",
 "report modal",
 "covoiturage",
 "alimentation durable",
 "engrais minéraux azotés",
 "sortie du pétrole"
 )

```

## climate subset building
```{r climate subset}
# Apply the specified terms to speech texts and create the corresponding match indicator or subset. Verify the unit of analysis before interpreting the resulting counts.

climate_pattern <- regex(
 paste0("\\b(", paste(climate_keywords, collapse = "|"), ")\\b"), ignore_case = TRUE
)

df_climate <- df_full %>%
 filter(str_detect(speech_text, climate_pattern))

dim(df_climate)
nrow(df_climate) / nrow(df_full)

df_climate %>%
 count(period) %>%
 mutate(ratio = n / sum(n))
```

## sample texts inspection
```{r text inspection}
# Inspect intermediate results for coverage, plausible matches and potential noise. This is a diagnostic step rather than a separate inferential test.

set.seed(123)

sample_climate_policy <- df_climate %>%
 slice_sample(n = min(5, nrow(df_climate))) %>%
 mutate(id = row_number())

textfile <- capture.output({
for(i in seq_len(nrow(sample_climate_policy))) {
 cat("\n\n=====================\n")
 cat("ID:", sample_climate_policy$id[i],"\n")
 cat("YEAR:", sample_climate_policy$year[i],"\n")
 cat("PERIOD:", sample_climate_policy$period[i], "\n\n")
 cat(sample_climate_policy$speech_text[i])
 cat("\n")
 }
})

writeLines(textfile, "rev_climate_policy_manual_inspection.txt")
```

## matched keywords check
```{r keywords matching, results='hide'}
# Define, combine or inspect the keyword set used in the subsequent matching step. Keyword selection determines the breadth of the resulting indicator.

sample_text <- sample_climate_policy %>%
 filter(id == 5) %>%
 pull(speech_text)

matches <- sapply(climate_keywords, function(keyword) {
 str_detect(sample_text, regex(paste0("\\b", keyword, "\\b"), ignore_case = TRUE))
})

matched_phrases <- climate_keywords[matches]

matched_phrases

data.frame(
 keyword = climate_keywords,
 matched = matches
) %>%
 filter(matched == TRUE)

```

## CCC explicit dictionary
```{r CCC explicit dictionary}
# Define, combine or inspect the keyword set used in the subsequent matching step. Keyword selection determines the breadth of the resulting indicator.

ccc_explicit <- c(
 "convention citoyenne",
 "convention citoyenne pour le climat",
 "convention citoyenne sur le climat",
 "conventions citoyennes",
 "citoyens de la convention",
 "citoyens tirés au sort",
 "tirage au sort",
 "150 citoyens"
 )

pattern_ccc_explicit <- paste(ccc_explicit, collapse = "|")
```

## CCC explicit keywords subset 
```{r CCC explicit keywords subset}
# Define, combine or inspect the keyword set used in the subsequent matching step. Keyword selection determines the breadth of the resulting indicator.

explicit_ccc_pattern <- regex(
 paste0(
  "\\b(",
  paste(ccc_explicit, collapse = "|"),
  ")\\b"
 ),
 ignore_case = TRUE
)

df_ccc_explicit <- df_climate %>%
 filter(
  str_detect(
   speech_text,
   explicit_ccc_pattern
  )
 )

dim(df_ccc_explicit)
nrow(df_ccc_explicit) / nrow(df_climate)
table(df_ccc_explicit$period, useNA = "ifany")

df_ccc_explicit %>%
 count(period) %>%
 mutate(ratio = n / sum(n))
```

## CCC explicit results check
```{r results check_explicit CCC, results='hide'}
# Inspect intermediate results for coverage, plausible matches and potential noise. This is a diagnostic step rather than a separate inferential test.

set.seed(123)

sample_ccc_explicit <- df_ccc_explicit %>%
 slice_sample(n = min(10, nrow(df_ccc_explicit))) %>%
 mutate(id = row_number())

textfile <- capture.output({
for(i in seq_len(nrow(sample_ccc_explicit))) {
 cat("\n\n=====================\n")
 cat("ID:", sample_ccc_explicit$id[i],"\n")
 cat("YEAR:", sample_ccc_explicit$year[i],"\n")
 cat("PERIOD:", sample_ccc_explicit$period[i], "\n\n")
 cat(sample_ccc_explicit$speech_text[i])
 cat("\n")
 }
})

writeLines(textfile, "rev_ccc_explicit_manual_inspection.txt")
```


## CCC policy cluster keywords: (1) consumption - publicité, recyclage removed
```{r CCC policy cluster_consumption}
# Define or inspect the consumption-cluster keyword set. The subsequent matching step treats these terms as indicators of thematic overlap, not proof of an explicit CCC reference.

consumption <- c(
  "surconsommation",
  "emballage",
  "plastique à usage unique", 
  "suremballage", 
  "consommation responsable", 
  "affichage environnemental",
  "affichage carbone",
  "écoresponsables",
  "écoresponsabilité",
  "produits polluants",
  "produits vertueux", 
  "économie circulaire",
  "déchets d'emballage",
  "utilisation du plastique",
  "vrac",
  "réutilisable",
  "contenants réutilisables",
  "bouteilles plastiques",
  "réduction des déchets",
  "matériaux biodégradables",
  "réemploi"
)

```

## CCC policy cluster keywords: (2) Production & Work
```{r CCC policy cluster_production_work}
# Define or inspect production-and-work terms in the policy-theme dictionary. Review broad terms for context-dependent matches.

production_work <- c(
  "produits durables",
  "écoconception",
  "longévité des produits", 
  "investissements verts",
  "clauses environnementales",
  "droit pollueur-payeur", 
  "filières émettrices", 
  "écocide",
  "produits micropolluants",
  "énergie verte",
  "énergies renouvelables",
  "méthanisation",
  "autoconsommation",
  "écolabels",
  "énergie décarbonée"
)
```

## CCC policy cluster keywords: (3) Transport & Mobility
```{r CCC policy cluster_transport_mobility}
# Define or inspect the transport-and-mobility keyword set used to identify relevant speech texts.

transport_mobility <- c(
  "mobilité durable",
  "véhicules faibles émissions",
  "mobilités douces",
  "véhicules partagés",
  "mobilités alternatives",
  "zones de faible émission",
  "pistes cyclables",
  "véhicules électriques",
  "véhicules hybrides",
  "émissions du véhicule",
  "écoconduite",
  "multimodalité",
  "moteur éthanol",
  "moteur électrique",
  "moteur thermique",
  "covoiturage",
  "biocarburants"
)

```

## CCC policy cluster keywords: (4) Housing - climatisation removed
```{r CCC policy cluster_housing}
# Define or inspect housing-related keywords and their matching coverage.

housing <- c(
  "construction bas carbone",
  "performance énergétique", 
  "rénovation globale",
  "chaudières au fioul",
  "passoires énergétiques",
  "bâtiments énergivores", 
  "audit énergétique",
  "certificats d'économies d'énergie", 
  "déperditions de chaleur",
  "chauffage au fioul"
)
```

## CCC policy cluster keywords: (5) Food
```{r CCC policy cluster_food}
# Define or inspect the food-policy cluster terms used in the thematic matching procedure.

food <- c(
  "alimentaire durable",
  "EGalim",
  "repas végétariens",
  "restauration collective",
  "faible coût environnemental",
  "transition vers le bio",
  "gaspillage alimentaire",
  "pratiques agroécologiques",
  "agroécologie",
  "réduction des pesticides",
  "agriculture biologique",
  "plan protéagineux",
  "prairies permanentes",
  "reconversion de l'élevage",
  "développement des légumineuses",
  "rediversifier les cultures",
  "engrais chimiques",
  "couvert végétal",
  "élevages de ruminants",
  "ecoscheme",
  "pêche durable",
  "surpêche",
  "propulsion verte",
  "clauses environnementales",
  "consommation de viande",
  "régime moins carné"
)

```

## CCC policy cluster keywords: (6) Constitution & Governance
```{r CCC policy cluster_constitution_governance}
# Define or inspect the constitution-and-governance cluster terms. Interpret its counts in light of the keyword set's narrower coverage.

constitution_gov <- c(
  "crime d'écocide",
  "érosion de la biodiversité",
  "utilisation des sols",
  "pollution chimique",
  "dommage écologique",
  "préjudice écologique"
)

```

## Merging CCC policy-theme keywords  
```{r ccc policy-theme merged}
# The keywords by cluster were merged into a single data frame to streamline to streamline the data wrangling process.

ccc_policy <- c(
 consumption,
 production_work,
 transport_mobility,
 housing,
 food,
 constitution_gov
)
ccc_policy <- unique(ccc_policy)

```

## CCC policy-theme subset building
```{r CCC policy-theme subset}
# Identify speeches containing one or more CCC policy-theme terms. Such matches indicate thematic overlap and do not necessarily mention the CCC.

ccc_policy_pattern <- regex(
 paste0(
  "\\b(",
  paste(ccc_policy, collapse = "|"),
  ")\\b"),
ignore_case = TRUE)

df_ccc_policy <- df_climate %>%
 filter(
  str_detect(
   speech_text,
   ccc_policy_pattern
  )
 )

  
dim(df_ccc_policy)
nrow(df_ccc_policy) / nrow(df_climate)
table(df_ccc_policy$period)
```

## sailience comparison
```{r salience comparison}
# Carry out the 'sailience comparison' step and inspect its output before proceeding. This chunk forms part of the documented workflow; its results have not been independently re-run here.

df_ccc_policy %>%
 count(period) %>%
 left_join(df_climate %>% count(period), by = "period",
           suffix = c("_ccc", "_climate")) %>%
 mutate(ratio = n_ccc / n_climate)
```

## create text file for manual inspection
```{r text file creation}
# Inspect intermediate results for coverage, plausible matches and potential noise. This is a diagnostic step rather than a separate inferential test.

set.seed(5678)

sample_ccc_policy <- df_ccc_policy %>%
 group_by(period) %>%
 slice_sample(n = 10) %>%
 ungroup() %>%
 arrange(period) %>%
 mutate(id = row_number())

textfile <- capture.output({
for(i in seq_len(nrow(sample_ccc_policy))) {
 cat("\n\n=====================\n")
 cat("ID:", sample_ccc_policy$id[i],"\n")
 cat("YEAR:", sample_ccc_policy$year[i],"\n")
 cat("PERIOD:", sample_ccc_policy$period[i], "\n\n")
 cat(sample_ccc_policy$speech_text[i])
 cat("\n")
 }
})

writeLines(textfile, "rbst_ccc_policy_manual_inspection.txt")

```

## matched phrase check
```{r matched phrase check, results='hide'}
# Inspect intermediate results for coverage, plausible matches and potential noise. This is a diagnostic step rather than a separate inferential test.

sample_text <- sample_ccc_policy %>%
 filter(id == 10) %>%
 pull(speech_text)

matches <- sapply(ccc_policy, function(keyword) {
 str_detect(sample_text, regex(paste0("\\b", keyword, "\\b"), ignore_case = TRUE))
})

matched_phrases <- ccc_policy[matches]

matched_phrases

data.frame(
 keyword = ccc_policy,
 matched = matches
) %>%
 filter(matched == TRUE)

```

## cluster keyword lists
```{r cluster keyword lists}
# Define, combine or inspect the keyword set used in the subsequent matching step. Keyword selection determines the breadth of the resulting indicator.

consumption_keywords <- c(
 "surconsommation",
  "emballage",
  "plastique à usage unique", 
  "suremballage", 
  "consommation responsable", 
  "affichage environnemental",
  "affichage carbone",
  "écoresponsables",
  "écoresponsabilité",
  "produits polluants",
  "produits vertueux", 
  "économie circulaire",
  "déchets d'emballage",
  "utilisation du plastique",
  "vrac",
  "réutilisable",
  "contenants réutilisables",
  "bouteilles plastiques",
  "réduction des déchets",
  "matériaux biodégradables",
  "réemploi")

production_keywords <- c(
 "produits durables",
  "écoconception",
  "longévité des produits", 
  "investissements verts",
  "clauses environnementales",
  "droit pollueur-payeur", 
  "filières émettrices", 
  "écocide",
  "produits micropolluants",
  "énergie verte",
  "énergies renouvelables",
  "méthanisation",
  "autoconsommation",
  "écolabels",
  "énergie décarbonée"
)

mobility_keywords <- c(
 "mobilité durable",
  "véhicules faibles émissions",
  "mobilités douces",
  "véhicules partagés",
  "mobilités alternatives",
  "zones de faible émission",
  "pistes cyclables",
  "véhicules électriques",
  "véhicules hybrides",
  "émissions du véhicule",
  "écoconduite",
  "multimodalité",
  "moteur éthanol",
  "moteur électrique",
  "moteur thermique",
  "covoiturage",
  "biocarburants"
)

housing_keywords <- c(
 "construction bas carbone",
  "performance énergétique", 
  "rénovation globale",
  "chaudières au fioul",
  "passoires énergétiques",
  "bâtiments énergivores", 
  "audit énergétique",
  "certificats d'économies d'énergie", 
  "déperditions de chaleur",
  "chauffage au fioul"
)

food_keywords <- c(
 "alimentaire durable",
  "EGalim",
  "repas végétariens",
  "restauration collective",
  "faible coût environnemental",
  "transition vers le bio",
  "gaspillage alimentaire",
  "pratiques agroécologiques",
  "agroécologie",
  "réduction des pesticides",
  "agriculture biologique",
  "plan protéagineux",
  "prairies permanentes",
  "reconversion de l'élevage",
  "développement des légumineuses",
  "rediversifier les cultures",
  "engrais chimiques",
  "couvert végétal",
  "élevages de ruminants",
  "ecoscheme",
  "pêche durable",
  "surpêche",
  "propulsion verte",
  "clauses environnementales",
  "consommation de viande",
  "régime moins carné"
)

constitution_keywords <- c(
 "crime d'écocide",
  "érosion de la biodiversité",
  "utilisation des sols",
  "pollution chimique",
  "dommage écologique",
  "préjudice écologique"
)
```

## pattern
```{r pattern}
# Carry out the 'pattern' step and inspect its output before proceeding. This chunk forms part of the documented workflow; its results have not been independently re-run here.

pattern <- function(keywords) {
 str_c("\\b(", str_c(keywords, collapse = "|"), ")\\b")
}

consumption_pattern <-
 pattern(consumption_keywords)
production_pattern <-
 pattern(production_keywords)
mobility_pattern <-
 pattern(mobility_keywords)
housing_pattern <-
 pattern(housing_keywords)
food_pattern <-
 pattern(food_keywords)
constitution_pattern <-
 pattern(constitution_keywords)

```

## creating cluster match variables
```{r variables for cluster matching}
# Apply the specified terms to speech texts and create the corresponding match indicator or subset. Verify the unit of analysis before interpreting the resulting counts.

data_cluster <- df_climate %>%
 mutate(
  text_lower = str_to_lower(speech_text),
  cluster_consumption =
   str_detect(text_lower,regex(consumption_pattern, ignore_case = TRUE)),
  cluster_production =
   str_detect(text_lower,regex(production_pattern, ignore_case = TRUE)),
  cluster_mobility =
   str_detect(text_lower, regex(mobility_pattern, ignore_case = TRUE)),
  cluster_housing =
   str_detect(text_lower, regex(housing_pattern, ignore_case = TRUE)),
  cluster_food =
   str_detect(text_lower, regex(food_pattern, ignore_case = TRUE)),
  cluster_constitution =
   str_detect(text_lower, regex(constitution_pattern, ignore_case = TRUE))
 )

cluster_long <- data_cluster %>%
 select(
  year, period, speech_text,
  cluster_consumption,
  cluster_production,
  cluster_mobility,
  cluster_housing,
  cluster_food,
  cluster_constitution) %>%
 pivot_longer(
  cols = starts_with("cluster_"),
  names_to = "cluster",
  values_to = "match"
 ) %>%
 filter(match == TRUE) %>%
 mutate(
  cluster = str_remove(cluster, "cluster_")
 )

```

## raw frequency
```{r raw frequency}
# Summarise matched speech counts for the specified categories. Raw counts should be interpreted alongside the corresponding corpus sizes.

cluster_raw <- cluster_long %>%
 count(cluster, period, name = "raw_count")

view(cluster_raw)
```

## periodic ratio
```{r periodic ratio}
# Calculate or display the relevant period-level proportions. The interpretation depends on the numerator, denominator and dictionary specification used in this chunk.

cluster_prop <- cluster_raw %>%
 group_by(period) %>%
 mutate(
  period_total = sum(raw_count),
  proportion = raw_count / period_total
 ) %>%
 ungroup()

cluster_prop <- cluster_prop %>%
 mutate(
  cluster = factor(
   cluster,
   levels = c(
    "consumption",
    "production",
    "mobility",
    "housing",
    "food",
    "constitution"
   )
  ),
   period = factor(
    period,
    levels = c(
     "pre",
     "transition",
     "post"
    )
   )
 ) %>%
 arrange(cluster, period) %>%
 mutate(
  proportion =
   round(proportion, 2)
 )

view(cluster_prop)

write.csv(cluster_prop, "cluster_table.csv", row.names = FALSE)


cluster_table <- cluster_prop %>%
 select(cluster, period, raw_count, proportion) %>%
 pivot_wider(
  names_from = period,
  values_from = c(raw_count, proportion)) %>%
   relocate(raw_count_pre:raw_count_transition, .before = raw_count_post)

cluster_table <- cluster_table %>%
 mutate(
  cluster = factor(
   cluster,
   levels = c(
    "consumption",
    "production",
    "mobility",
    "housing",
    "food",
    "constitution"
   )
  )
 ) %>%
 arrange(cluster) %>%
 mutate(
  proportion_pre =
   round(proportion_pre, 2),
  proportion_transition =
   round(proportion_transition, 2),
  proportion_post =
   round(proportion_post, 2)
 ) %>%
 select(
  cluster,
  raw_count_pre,
  proportion_pre,
  raw_count_transition,
  proportion_transition,
  raw_count_post,
  proportion_post
 )

view(cluster_table)
```

## visualisation

```{r visualisation}
# Produce the figure for the preceding analysis. Check the axis labels, denominator and time grouping against the data object used in this chunk.

ggplot(cluster_prop, mapping = aes(x = period, y = proportion, group = cluster)) +
 geom_line() +
 geom_point() +
 facet_wrap(~ cluster, scales = "free_y") +
 labs(
  title = "Cluster-level robustness check",
  x = "Period",
  y = "Proportion within CCC policy-theme matches"
 ) +
 theme_minimal()

ggsave(
  "cluster_robustness.pdf",
  width = 10,
  height = 7
)
```



