---
title: "European Climate Politics: Party Positions and Trends"
author: "Junhui Park"
date: "`r Sys.Date()`"
output: html_document
---

# 1. Introduction
Climate change has become one of the most contested political issues in Western Europe. This project examines how political parties in four countries (France, Germany, Sweden, and the United Kingdom) differ in their positions towards environmental protection, using the Comparative Manifesto Project (CMP) dataset obtained from the WZB Berlin Social Sciences Centre (https://manifesto-project.wzb.eu).
This analysis focuses on per501 (Environmental Protection) in domain 5 (Welfare and Quality of LIfe) of the dataset.
```{r, message=FALSE, warning=FALSE, include=FALSE}
library(manifestoR)
library(tidyverse)
mp_setapikey(key = "af74549df67d5fb1f281d60a9158c579")
```

# 2. Data
```{r, message=FALSE, warning=FALSE, results='hide'}
cmp <- mp_maindataset()

country <- cmp %>%
 filter(countryname %in% c("Germany", "France", "Sweden", "United Kingdom")) %>%
 mutate(year = date %/% 100)
```


# 3. Descriptive Statistics
```{r}
# key variable overview
country %>%
 select(countryname, partyname, parfam, year, pervote, rile, per501) %>%
 glimpse()

# number of observation by country
country %>%
 count(countryname)

# summary statistics of environmental protection variable (per501)
summary(country$per501)
```

# 4. Visualisation
```{r}
library(ggplot2)

# time-series data by country

country_mean <- country %>%
 group_by(countryname, year) %>%
 summarise(mean_per501 = mean(per501)) %>%
 ungroup()

 
 ggplot(country_mean, aes(x = year, y = mean_per501, colour = countryname)) +
 geom_line() +
  labs(title = "Time-series data of the environmental protection manifesto by country",
       x = "Year",
       y = "Average of per501",
       color = "Country") +
  theme_minimal()

# time-series data by party family

party_family <- country %>%
 mutate(parfam_label = case_when(
  parfam == 10 ~ "ECO",
  parfam == 30 ~ "SOC",
  parfam == 50 ~ "CHR",
  parfam == 60 ~ "CON",
  parfam == 70 ~ "NAT",
  TRUE ~ "Other"
 )) %>%
 filter(parfam_label != "Other")

 
ggplot(party_family,
       aes(x = year, y = per501, colour = parfam_label)) +
 geom_line() +
 facet_wrap(~ countryname) +
 labs(
  title = "Time-series data of the environmental protection manifesto by party family",
  x = "Year",
  y = "Environmental protection manifesto",
  color = "Party family"
 ) +
 theme_minimal()

```

# 5. Regression Analysis
```{r, results='asis'}
model1 <- lm(per501 ~ year, country)
summary(model1)

model2 <- lm(per501 ~ year + parfam_label, party_family)
summary(model2)

model3 <- lm(per501 ~ year + parfam_label + countryname, party_family)
summary(model3)

library(stargazer)
stargazer(model1, model2, model3, 
          type = "html",
          title = "Regression Results: Determinants of Environmental Protection Manifesto",
          dep.var.labels = "Environmental Protection (per501)",
          covariate.labels = c("Year", "CON", "ECO", "NAT", "SOC", 
                               "Germany", "Sweden", "United Kingdom"))

```
# 6. Discussion
The average environmental protection manifesto scores across all four countries display a similar fluctuating pattern over time. The aggregated figures for the United Kingdom remain comparetively modest throughout the observed period. Germany and Sweden both reach their peak in the late 1980s, while France records its highest level somewhat later, in the late 1990s. Most notably, Sweden consistently shows the highest scores from the late 1970s onward. In addition, manifesto scores across all countries show a declining trend following 2020.

The time-series data by party family provides a more fine-grained picture. In Western Europe, Social Democratic and Christian Democratic parties have historically dominated party systems. Given the subsequent emergence of Green parties and radical right parties, this analysis focuses on five party families: 1) CHR: Christian democratic parties, 2) CON: Conservative parties, 3) ECO: Ecological parties, 4) NAT: Nationalist and radical right parties, and 5) SOC: Social democratic parties. With the exception of the United Kingdom, environmental protection manifesto scores begin to emerge between the 1980s and 1990s. Sweden reaches its peak in the late 1980s, followed by a substantial decline. Notably, in Germany, high manifesto scores around the same period are observed across all three major party families - CHR, SOC, and ECO, suggesting a broader contagion effect beyond Green parties alone. By contrast, ECO parties' environmental protection scores accounts for a comparatively smaller share between the late 2010s and 2020s.

These figures suggest that Green parties emerged earlier in Sweden and secured more prominent positions within the party system compared to other countries. The stronger the presence of Green parties in party systems, the higher the overall environmental protection manifesto scores tend to be. The results of a multiple OLS regression analysis are broadly consistent with these pattern. When controlling for party family and country, the coefficients of ECO party family and Sweden are statistically significant. 

However, this analysis is not without limitation. The variable selection, limited to country and party family, captures only a small portion of variation in environmental protection manifesto scores. Furthermore, high scores among Green parties do not necessarily imply that other party families neglect environmental issues in their programmes. Party competition dynamics warrant further consideration. A substantial body of research on party competition has demonstrated that mainstream and challenger parties tend to mobilise issues outside their traditional agendas when facing electoral threats or opportunities, a dynamic that may also shape environmental manifesto commitments.